Canonical rings of $\bq$-divisors on Minimal Rational Surfaces

### 2015-07-31 surfaces-write-up/surface-write-up.tex
confused about sigma

\end{lem}
\begin{proof}
Let 
\begin{align}\label{eqn:Sigma-def}
	\Sigma = \left \{(d, c_1, \ldots, c_n) \in \bz^{n+1} : c_i \geq - d
\alpha_i, \; 1 \leq i \leq n, \text{ and } \sum_{i=1}^{n} \ell_i \bida_i = 0
\right \}.
\end{align}
\todo{Aaron: Possibly take this definition out of the proof, and move it into
	the setup, since we want to reference it later.  \\
Peter: I like having the statement of the lemma being nice and consise; also
the statement has nothing to do with this geometric picture, so I find it more
natural to have the geometric side contained in the proof.  I also don't quite
see why it is so bad to reference something from inside a previous proof; a
reader can surely locate it...

Aaron: In general, I like the idea of having proofs self contained,
so that one should only refer to the proof statement, not the
body of the proof.
I'm not very strongly tied to this idea though, and I think it's fine. I think it's fine.
\\
Peter: I understand where your objection is coming from.  However, I don't see a good way to fix it without creating a very ugly statement of the lemma.
}
We note that $\Sigma$ has extremal rays given by the lattice points 
\begin{equation}\label{defn:e-i-proj}
	e_i = \left(\ell_i \bida_i, - \alpha_1 \ell_i \bida_i, \ldots
-\alpha_{i-1} \ell_i \bida_i, \ell_i \sum_{j\ne i} \alpha_j \bida_j,
-\alpha_{i+1} \ell_i \bida_i, \ldots, -\alpha_n, \ell_i \bida_i \right)
\end{equation}
for each $i\in \{1, \ldots n\}$.
Therefore, applying Proposition ~\ref{prop:cone-generation}, we see $R_D$ is generated in degrees less than
\[
	\sum_{i=1}^n \ell_i a_i.
\]
\end{proof}

@@

Peter: Better (on the first comment? As for the second, that might be a little tricky, because we already have $E_\mu$, so we would have to do something like $E_{i,\mu}$ (which we could of course do...)}
we simply write $f_i$ as a 
polynomial in $f_1, \ldots f_4$. Furthermore, these
generate all relations, since they allow us to reduce any $\prod_{j =
1}^n f_j^{c_j}$ to the canonical form where $r_i = - \lfloor d \alpha_i
\rfloor$ whenever $i > 4$.  To bound the degree of generation of these relations, we bound the degree of generation of the ideal $(z_i - \beta_i) \cap \ker(\psi)$ for each $i$.

@@

Let $\omega\in \Omega'$ correspond to the monomial $y_i \prod_{j=1}^n
{y_j}^{c_j}$.  Then we can write $\omega$ as a sum of $r_j e_j$'s for $j\in
\Te$ and $r_{j,k}e_{j,k}$ for $j\in \Tp, k\in \Tm$.  
For convenience, let $e_j$ (when it exists) be in degree $d_j$ with $j^{th}$
component $-\alpha_j d_j + \kappa_j$ for some $\kappa_j \in \bq$ and let $e_{j,k}$ (when it exists) lie in
degrees $d_{j,k}$ with $j^{th}$ component $-\alpha_j d_{j,k} + \kappa_{j,k}'$ and $k^{th}$
component $-\alpha_j d_{j,k} + \kappa_{j,k}''$ for $\kappa_{j,k}', \kappa_{j,k}'' \in \bq$.
By a careful computation, we can find $\lambda \in \Sigma$ 
and $r_{i,j} \in \bq$ (in general such a choice may be very non-unique) so that we can write
\begin{equation}\label{eqn:sigma-in-terms-of-lambda/E}
	\sigma = \lambda + E
 \end{equation}
with $E$ of the form
\begin{equation}\label{eqn:hirz-E-translation}
	E = \begin{cases}
	\frac{1}{\kappa_i} e_i  	&  \mbox{ if }i\in \Te\\
	\sum_{j\in \Tm} r_{i,j} e_{i,j}: r_{i,j}\in \bq, \sum_{j\in \Tm} r_{i,j}\kappa_{i,j}' = 1 & \mbox{ if } i\in \Tp\\
	\sum_{j\in \Tp} r_{j,i} e_{j,i}: r_{j,i} \in \bq, \sum_{j\in \Tp} r_{j,i} \kappa_{i,j}'' = 1 	& \mbox{ if } i\in \Tm\\
	\end{cases}.
\end{equation}
Specifically, if we write out $\sigma = \sum_{j\in \Te} s_j e_j + \sum_{j\in \Tp, k \in \Tp} s_{j,k} e_{j,k}$, then any choice of $r_{j,k}$'s not greater than the respective $s_{j,k}$'s such that $r_{i,j} \kappa_{i,j}' =1$ (when $i\in \Tp$ but a similar statement holds for $i\in \Tm$) will suffice.  Such a choice of $r_{i,j}$'s uniquely determines the necessary $\lambda$.

>>>>

\end{lem}
\begin{proof}
Let 
\begin{align}\label{eqn:Sigma-def}
	\Sigma = \left \{(d, c_1, \ldots, c_n) \in \bz^{n+1} : c_i \geq - d
\alpha_i, \; 1 \leq i \leq n, \text{ and } \sum_{i=1}^{n} \ell_i \bida_i = 0
\right \}.
\end{align}

We note that $\Sigma$ has extremal rays given by the lattice points 
\begin{equation}\label{defn:e-i-proj}
	e_i = \left(\ell_i \bida_i, - \alpha_1 \ell_i \bida_i, \ldots
-\alpha_{i-1} \ell_i \bida_i, \ell_i \sum_{j\ne i} \alpha_j \bida_j,
-\alpha_{i+1} \ell_i \bida_i, \ldots, -\alpha_n, \ell_i \bida_i \right)
\end{equation}
for each $i\in \{1, \ldots n\}$.
Therefore, applying Proposition ~\ref{prop:cone-generation}, we see $R_D$ is generated in degrees less than
\[
	\sum_{i=1}^n \ell_i a_i.
\]
\end{proof}

@@

Peter: Better (on the first comment? As for the second, that might be a little tricky, because we already have $E_\mu$, so we would have to do something like $E_{i,\mu}$ (which we could of course do...)

Aaron: Does $E_\mu$ depend on $i$? I thought it didn't, but if it does, then we should definitely do $E_{i,\mu}$ since I didn't
realize it depends on $i$. If $E_\mu$ doesn't depend on $i$, then we can just write $E_i$, and $E_\mu$,
and there is no danger of confusion because the subscripts live in different categories.}
we simply write $f_i$ as a 
polynomial in $f_1, \ldots f_4$. Furthermore, these
generate all relations, since they allow us to reduce any $\prod_{j =
1}^n f_j^{c_j}$ to the canonical form where $r_i = - \lfloor d \alpha_i
\rfloor$ whenever $i > 4$.  To bound the degree of generation of these relations, we bound the degree of generation of the ideal $(z_i - \beta_i) \cap \ker(\psi)$ for each $i$.

@@

Let $\omega\in \Omega'$ correspond to the monomial $y_i \prod_{j=1}^n
{y_j}^{c_j}$.  Then we can write $\omega$ as a sum of $r_j e_j$'s for $j\in
\Te$ and $r_{j,k}e_{j,k}$ for $j\in \Tp, k\in \Tm$.  
For convenience, let $e_j$ (when it exists) be in degree $d_j$ with $j^{th}$
component $-\alpha_j d_j + \kappa_j$ for some $\kappa_j \in \bq$ and let $e_{j,k}$ (when it exists) lie in
degrees $d_{j,k}$ with $j^{th}$ component $-\alpha_j d_{j,k} + \kappa_{j,k}'$ and $k^{th}$
component $-\alpha_j d_{j,k} + \kappa_{j,k}''$ for $\kappa_{j,k}', \kappa_{j,k}'' \in \bq$.
By a careful computation, we can find $\lambda \in \Sigma$ 
and $r_{i,j} \in \bq$ (in general such a choice may be very non-unique) so that we can write
\begin{equation}\label{eqn:sigma-in-terms-of-lambda/E}
	\sigma = \lambda + E
 \end{equation}

\todo{Is $\sigma = \omega$? If not I'm very confused, since I don't see $\sigma$ defined anywhere before or after
	but it seems like the key to the whole proof.
}
with $E$ of the form
\begin{equation}\label{eqn:hirz-E-translation}
	E = \begin{cases}
	\frac{1}{\kappa_i} e_i  	&  \mbox{ if }i\in \Te\\
	\sum_{j\in \Tm} r_{i,j} e_{i,j}: r_{i,j}\in \bq, \sum_{j\in \Tm} r_{i,j}\kappa_{i,j}' = 1 & \mbox{ if } i\in \Tp\\
	\sum_{j\in \Tp} r_{j,i} e_{j,i}: r_{j,i} \in \bq, \sum_{j\in \Tp} r_{j,i} \kappa_{i,j}'' = 1 	& \mbox{ if } i\in \Tm\\
	\end{cases}.
\end{equation}
Specifically, if we write out $\sigma = \sum_{j\in \Te} s_j e_j + \sum_{j\in \Tp, k \in \Tp} s_{j,k} e_{j,k}$, then any choice of $r_{j,k}$'s not greater than the respective $s_{j,k}$'s such that $r_{i,j} \kappa_{i,j}' =1$ (when $i\in \Tp$ but a similar statement holds for $i\in \Tm$) will suffice.  Such a choice of $r_{i,j}$'s uniquely determines the necessary $\lambda$.
